---
title: "taxnames: Formating taxonomic names in Markdown and LaTeX"
format:
  gfm:
    toc: true
---

# Motivation

This package attempts to simplify the citation and formatting of taxonomic names in Markdown documents ([R Markdown](https://rmarkdown.rstudio.com/) and [Quarto](https://quarto.org/)) as well as in LaTeX ([knitr](https://yihui.org/knitr/)).

To use this package, you need to have a taxonomic list imported into the [taxlist](https://docs.ropensci.org/taxlist/) package, including taxonomic ranks and parent-child relationships.

# Get started

## Installing taxnames

Assuming you are already working in one of the frameworks, either R Mardown or Quarto, you will need to install `taxnames` on your own system.

```{r}
#| eval: false
install.packages("taxnames", dependencies = TRUE)
```

Alternatively, you can download the latest development version from **GitHub** using the package [remotes](https://remotes.r-lib.org/).

```{r}
#| eval: false
library(remotes)
install_github("kamapu/taxnames")
```

## Set a standard taxonomy

Before you continue, you will need to load the packages `taxlist` and `taxnames` in your session.

```{r}
#| message: false
library(taxlist)
library(taxnames)
```

The package `taxnames` uses a taxonomic object formatted by `taxlist` and stored in the background.
If you manage to import your taxonomic list to `taxlist`, you can then set your own standard taxonomy.

```{r}
#| eval: false
set_tax(your_taxlist_object)
```

Per default `taxnames` takes the taxonomic list **Easplist** form the package `taxlist`.

```{r}
get_tax()
```